---
title: "EV Home and Work Assignment Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EV Home and Work Assignment Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we're going to assign 250,000 BEVs and 100,000 PHEVs to the study area and estimate total charging demand at commute destinations.

Load the required packages.
```{r setup}
library(tidyverse)
library(evworkplace)
```

The functions in this package have defaults set for all parameters, but it's a good idea to specify them individually. The main parameters for this analysis are:
* Maximum EV fraction in a single block group (default = 70%)
* Commute fraction for EVs (default = 60%)
* Extra miles to drive per day (default = 10)
```{r params}
max_ev_fraction  <- 0.7
commute_fraction <- 0.6
extra_miles      <- 10
```

Some other paramters can affect charging probabilities at homes and commute destinations.
The charging parameters for this analysis are:
* Fraction of EVs assigned to single family homes (default = 90%)
* Fraction of EVs in single family homes with access to charging at home (default = 90%)
* Fraction of EVs in multi-unit dwellings with access to charging at home (default = 10%)
* Is charging at commute destinations free to users? TRUE/FALSE
* Electricity rate paid for home charging (default = 12.97 cents per kWh)
* Proportion of long/medium/short range vehicles of each type to include (default = 50% long, 40% medium) ... range definitions for can be found in the documentation for the make_workcharge_probs function.
```{r params-charging}
single_family_home <- 0.9
sfh_home           <- 0.9
mud_home           <- 0.2
is_workcharge_free <- FALSE
rate_paid          <- 12.97
long_range         <- 0.5
medium_range       <- 0.4
short_range        <- 0.1
```



## Prepare data and assign EVs

First, we need to prepare the data. All of the input data is included in this package, but some cleaning is required to make the data usable.

First, let's look at the raw data, which has a range of information about each of the block groups in the study area. Note that `vehicles` provides the total number of household vehicles in the block group:
```{r}
dvrpc_bg_data
```

The vehicle assignment function requires input variables to be centered on 0 and scaled to have a standard deviation of 1. Additionally, this is where we specify the maximum potential number of vehicles that can be converted to EVs in any adoption scenario.
```{r prep-data}
prepped_data <- prep_bg_data(max_evs = max_ev_fraction)
prepped_data
```

Create and run vehicle assignment scenarios. Note that these scenarios are currently run independently, which has the benefit of making it possible to specify multiple BEV or PHEV scenarios at once. The downside of this is that the total number of vehicles assigned to a block group across a scenario each for BEVs and PHEVs may be greater than the total number of vehicles in the block group. This can generally be avoided by setting `max_ev_fraction` to 0.7 or lower and running scenarios with no more than 50% total vehicle electrification.
```{r assign-vehicles}
adoption_scenario <- prepped_data %>%
  assign_evs(250000, evworkplace::assignment_model_coefficients$CA_bev,  "bevs_250k") %>%
  assign_evs(100000, evworkplace::assignment_model_coefficients$CA_phev, "phevs_100k")

adoption_scenario <- adoption_scenario %>% 
  select(GEOID, bevs_250k, phevs_100k)

adoption_scenario

```

## Compute Demand for Charging at Commute Destinations

After assigning vehicles to households, the next step is to link those vehicles to commute destinations and estimate workplace charging needs.


First, assign the vehicles to commute destinations. This will likely take longer than previous steps. The only parameter to worry about here is the share of vehicles used for commuting. 
```{r assign-commute-dests}
commute_locations <- adoption_scenario %>% 
  prep_workcharge(bevs_250k, phevs_100k, .ev_commute_frac = commute_fraction)
```

Then, create a scenario for workplace charging probabilities. You can generate multiple scenarios to look at the impacts of charging parameters, but it's generally best to have at least one for BEVs and one for PHEVs. The function used here will generate home charging probabilities for vehicles with various commute distances.

In this example, a single set of parameters is used for both BEV and PHEV charging scenarios, but you could use different parameters for each vehicle type or generate multiple scenarios to compare. The `is_workcharge_free` variable has a particularly strong impact on people's charging behavior.
```{r make-probs}

bev_chargeprob  <- make_workcharge_probs(
  use_BEV = TRUE,
  single_family_home_frac = single_family_home,
  sfh_homecharge = sfh_home,
  mud_homecharge = mud_home,
  is_workcharge_free = is_workcharge_free,
  rate_paid = rate_paid,
  long_range_wt = long_range,
  medium_range_wt = medium_range,
  short_range_wt = short_range)

phev_chargeprob <- make_workcharge_probs(
  use_BEV = FALSE,
  single_family_home_frac = single_family_home,
  sfh_homecharge = sfh_home,
  mud_homecharge = mud_home,
  is_workcharge_free = is_workcharge_free,
  rate_paid = rate_paid,
  long_range_wt = long_range,
  medium_range_wt = medium_range,
  short_range_wt = short_range)


```

Finally, combine the commute assignment with the charging probabilities to create a complete set of charging probabilities at commute destinations.
```{r assign-workcharge}

ev_workplace_demand_bev250 <- commute_locations %>%
  assign_workcharge(bevs_250k, bev_chargeprob, "bev250k_basic")

ev_workplace_demand_phev100 <- commute_locations %>%
  assign_workcharge(phevs_100k, phev_chargeprob, "phev100k_basic")
```

## Save model outputs

You probably want to combine the adoption and workplace scenarios into a single table. You can do that like this:
```{r combine-tables}
output_table <- adoption_scenario %>% 
  left_join(ev_workplace_demand_bev250, by =  c("GEOID" = "Work_BlkGrp")) %>% 
  left_join(ev_workplace_demand_phev100, by = c("GEOID" = "Work_BlkGrp"))
```


You can save the adoption scenario and workplace charging results as a csv file like this:
```{r write-csv, eval=FALSE}
output_directory <- "my/directory/path" # replace this with a path to a folder where you'd like to store the results.

output_table %>% 
  write_csv(file.path(output_directory, "adoption_and_workcharging_scenario.csv"))
```

Alternatively, you can export the data directly to the clipboard and paste it into excel or another spreadsheet program. This can be especially useful if you're working in a state that has a FIPS code starting starting with "0" (AK, AL, AR, AZ, CA, CO, CT).
```{r write-clip, eval=FALSE}
output_table %>% 
  clipr::write_clip()

```

Alternatively, you can attach the columns to a shapefile of block group boundaries using the SF package (although you might need to adjust column names / lengths a bit to get it to write properly):
```{r write-shp, eval=FALSE}

library(sf)

# load a shapefile of block group boundaries (NOTE: the GEOID / FIPS column in this dataset must be formatted as character/text)
bg_shp <- read_sf("my/shapefile/directory/dvrpc_bgs.shp")

shp_fips_colname <- "GEOID"

# attach adoption and charging details to this shapefile
bg_shp_with_details <- bg_shp %>% 
  left_join(output_table, by = c(shp_fips_colname = "GEOID"))

# and write to disk
bg_shp_with_details %>% 
  write_sf("my/shapefile/directory/dvrpc_bgs_evscenario.shp")

```

